TITLE by Andre Kenji Yai
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

#install.packages('GGally')
#install.packages('scales')
#install.packages('lattice')
#install.packages('MASS')
#install.packages('car')
#install.packages('reshape')
#install.packages('plyr')

library(ggplot2)
library(GGally)
library(scales)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
WhiteWines <- read.csv('wineQualityWhites.csv')
```

> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

Let's start our analysis by summarizing the data and getting to know more about the dataset.

```{r echo=FALSE, Univariate_Plots}
dim(WhiteWines)
```

In our dataset there are 4898 rows and 13 features and the features are:

```{r}
names(WhiteWines)
```

Let's see the first rows of the dataset

```{r}
head(WhiteWines,n=10)
```

We got that our features are numerical and most are double and it seems X the id of the row.

Lets summarize the data to know more about the mean and perticiles of each feature.
```{r}
summary(WhiteWines)
```


There are two categories of variables in this database:

- continuous: ph, alcohol, sulphates, clorides, free sulfur dioxide, total sulfur dioxide, density, residual sugar and acids,
- descrite: X, quality.

We also noticed that those with higher variance are alcohol, free sulfur dioxide, total sulfur dioxide, residual sugar. 

We will start analysing the quality that and those variables with high variance

#### Quality

Lets see how the wines were ranked.

```{r}

WineQuality <- table(WhiteWines$quality)
barplot(WineQuality, main="Wine Quality Distribution", xlab="Number of Wines")
```

We got that Wine Quality distribution look like a Normal Distribution with the most wines was ranked with a 6 followed by 5 and the lowest received a 3 and the highest a 9.

I wonder what can contribute to that grades? So lets look now at those with higher variance and how they are distributed. 

### Alcohol

The porcentage of alcohol content of the wine.

```{r}
qplot(x = alcohol, data = WhiteWines)
```


```{r}
summary(WhiteWines$alcohol)
```

We got a left skewed histogram with the most around 10 lowest a 8.00 and the highest at 14.20.
I wonder if those with more than 10 have a highest ranking of quality.

### Residual sugar

The amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

```{r}
qplot(x = residual.sugar, data = WhiteWines)
```
```{r}
summary(WhiteWines$residual.sugar)
```

```{r}
length(WhiteWines[(WhiteWines$residual.sugar < 1),]$residual.sugar)
length(WhiteWines[(WhiteWines$residual.sugar > 45),]$residual.sugar)
```

We also got a left skewed distribution with most of the data concentrated in less than 20 g/L. With the less at 0.60 and highest at 65.800 g/L and second highest at 31.60 g/L.

That been investigated we got that we have 77 less than 1 and 1 wine with residual sugar more than 45 g/L.

### total sulfur dioxide

Amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

```{r}
qplot(x = total.sulfur.dioxide, data = WhiteWines)
```

```{r}
summary(WhiteWines$total.sulfur.dioxide)
```

It is a normal with it peak near 130 ppm, most of the data concentrated between 0-250 ppm and the minimum at 9 ppm and max at 440ppm.


```{r}
length(WhiteWines[(WhiteWines$total.sulfur.dioxide > 50),]$total.sulfur.dioxide)/length(WhiteWines$total.sulfur.dioxide)
```

We also got that around 98.9% of the dataset have a more the SO2 more present in the wine. 

# free sulfur dioxide

The free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine.

```{r}
qplot(x = free.sulfur.dioxide, data = WhiteWines)
```

```{r}
summary(WhiteWines$free.sulfur.dioxide)
```

Also a normal with concentration in beetween 0 and 100 and it peak near 50 and min value 
at 2 and max 289.


Lets see the others values as well. I suppose that ph and density may be also correlated.

#### pH

Describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale


```{r}
qplot(x=pH, data=WhiteWines)
```

```{r}
summary(WhiteWines$pH)
```


We get that all wines are acid (pH < 7) and it has a normal distribution most of it at 3.1
and min 2.720 and max 3.820.

#### Density

The density of water is close to that of water depending on the percent alcohol and sugar content.

```{r}
qplot(x=density,data=WhiteWines)
```


```{r}
summary(WhiteWines$density)
```

It is a normal peak at of it at 0.9937 and min at 0.987 and max near 1.04. It is concentrated at 0.9 to 1.00. That means that density of most wines are very near to the water density, that is 1.

## Chlorides

Is the amount of salt in the wine. I wonder how this can affect quality.

```{r}
qplot(x=chlorides,data=WhiteWines)
```


```{r}
summary(WhiteWines$chlorides)
```

Is also a normal with peak at 0.043 and it is concentrated between  0.009 and 0,09 and the min is at 0.009 and max at 0.34. I suppose  be the less than 0.04 have best quality,


#### Sulphates

A wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant


```{r}
qplot(x=sulphates,data=WhiteWines)
```


```{r}
summary(WhiteWines$sulphates)
```

We got a binormal with peaks near 0.4 and 0.5. Min value is 0.2 and max 1.08.


#### Acids

Lwts see the acid. There are three types of acids:
   1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
   2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
    3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
    
    
```{r}
qplot(x=fixed.acidity,data=WhiteWines)
```

```{r}
summary(WhiteWines$fixed.acidity)
```


```{r}
qplot(x=volatile.acidity,data=WhiteWines)
```
```{r}
summary(WhiteWines$volatile.acidity)
```

```{r}
qplot(x=citric.acid,data=WhiteWines)
```

```{r}
summary(WhiteWines$citric.acid)
```

Comparasion:

In all of them we got a normal. We have more fixed acid than the others and it peak is in 7 when the other is in 0.3 and with a higher variance.


Lets aggregate the acids and see what it returns.

```{r}
WhiteWines$acid <- WhiteWines$citric.acid + WhiteWines$fixed.acidity + WhiteWines$volatile.acidity
```

```{r}
qplot(x=acid,data=WhiteWines)
```

```{r}
summary(WhiteWines$acid)
```






> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

We will now see the correlation matrix. To see wich variables are correlated to quality and each other.

```{r}
correlation_matrix <- cor(WhiteWines)
round(correlation_matrix,2)
```

Looking more carefully with the row quality we got that quality is positivly and more correlated to alcohol. Other feature that is correlated but negatively is with density and clorides.

```{r}
sort(correlation_matrix[13,])
```

Lets  see a visualization of this matrix.

```{r echo=FALSE, Bivariate_Plots}
library(psych)
pairs.panels(WhiteWines, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.


Lets take a closer look at the relation between alcohol and density with quality.

```{r}
WhiteWines$Qualitycat<-as.factor(WhiteWines$quality)
```


```{r}
ggplot(aes(y = alcohol,x = Qualitycat),data = WhiteWines) + geom_boxplot() + geom_smooth(method = 'lm')
```

```{r}
tapply(WhiteWines$alcohol,WhiteWines$Qualitycat, summary)
```

We see that as alcohol in the wines increases the quality tend to increase as well and that with alcohol percentage more than 10.8 tend to have better quality.
Now lets take a look at density.

```{r}
ggplot(aes(x = Qualitycat, y = density),data = WhiteWines) + geom_boxplot(color = '#F79420')
```


```{r}
tapply(WhiteWines$density,WhiteWines$Qualitycat, summary)
```

It seems that low density tend to a better quality. As it goes to near 0.99 it's Quality tend to increases.

Now let us compare to chorlides.


```{r}
ggplot(aes(x = Qualitycat, y = chlorides),data = WhiteWines) + geom_boxplot(color = '#F79420')
```






### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!