White Wines Quality Analysis by Andre Kenji Yai
========================================================

```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

#install.packages('GGally')
#install.packages('scales')
#install.packages('lattice')
#install.packages('MASS')
#install.packages('car')
#install.packages('reshape')
#install.packages('plyr')

library(ggplot2)
library(GGally)
library(scales)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
WhiteWines <- read.csv('wineQualityWhites.csv')
```

# Description 

  P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
  Modeling wine preferences by data mining from physicochemical properties.
  In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

  In the above reference, two datasets were created, using red and white wine samples.
  The inputs include objective tests (e.g. PH values) and the output is based on sensory data
  (median of at least 3 evaluations made by wine experts). Each expert graded the wine quality 
  between 0 (very bad) and 10 (very excellent). Several data mining methods were applied to model
  these datasets under a regression approach. The support vector machine model achieved the
  best results. Several metrics were computed: MAD, confusion matrix for a fixed error tolerance (T),
  etc. Also, we plot the relative importances of the input variables (as measured by a sensitivity  analysis procedure).
  
  Number of Attributes: 11 + output attribute
  
   Note: several of the attributes may be correlated, thus it makes sense to apply some sort of
   feature selection.

  Attribute information:

   For more information, read [Cortez et al., 2009].

   Input variables (based on physicochemical tests):
    - 1.  fixed acidity (tartaric acid - g / dm^3)
    - 2. volatile acidity (acetic acid - g / dm^3)
    - 3. citric acid (g / dm^3)
    - 4. residual sugar (g / dm^3)
    - 5. chlorides (sodium chloride - g / dm^3
    - 6. free sulfur dioxide (mg / dm^3)
    - 7. total sulfur dioxide (mg / dm^3)
    - 8. density (g / cm^3)
    - 9. pH
    - 10. sulphates (potassium sulphate - g / dm3)
    - 11. alcohol (% by volume)
    Output variable (based on sensory data): 
    - 12. quality (score between 0 and 10)

# The Problem

  We will analiyze the dataset to investigate the features that makes a good wine. In this problem we will use the White Wine dataset. 

# Univariate Plots Section

In this section, I will perform some preliminary exploration of your dataset.

Let's start our analysis by summarizing the data and getting to know more about the dataset.

```{r echo=FALSE, Univariate_Plots}
dim(WhiteWines)
```

In our dataset there are 4898 rows and 13 features and the features are:

```{r}
names(WhiteWines)
```

Let's see the first rows of the dataset

```{r}
head(WhiteWines,n=10)
```

We got that our features are numerical and most are double and it seems X the id of the row.

Lets summarize the data to know more about the mean and perticiles of each feature.
```{r}
summary(WhiteWines)
```


There are two categories of variables in this database:

- continuous: ph, alcohol, sulphates, clorides, free sulfur dioxide, total sulfur dioxide, density, residual sugar and acids,
- descrite: X, quality.

We also noticed that those with higher variance are alcohol, free sulfur dioxide, total sulfur dioxide, residual sugar. 

We will start analysing the quality that and those variables with high variance


### Quality

Lets see how the wines were ranked.

```{r}
WineQuality <- table(WhiteWines$quality)
barplot(WineQuality, main="Wine Quality Distribution", xlab="Number of Wines")
```


We got that Wine Quality distribution look like a Normal Distribution with the most wines was ranked with a 6 followed by 5 and the lowest received a 3 and the highest a 9.

I wonder what can contribute to that grades? So lets look now at those with higher variance and how they are distributed. 

As we have a descrite and specific set of values assigned to  quality, I will factorize this feature. Doing so it will help us in the visualization by allowing us to perform better boxplots and investigate relationship of features.

```{r}
WhiteWines$QualityCategory <-as.factor(WhiteWines$quality)
```



### Alcohol

The porcentage of alcohol content of the wine.

```{r}
qplot(x = alcohol, data = WhiteWines)
```


```{r}
summary(WhiteWines$alcohol)
```

We got a left skewed histogram with the most around 10 lowest a 8.00 and the highest at 14.20.
I wonder if those with more than 10 have a highest ranking of quality.

### Residual Sugar

The amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

```{r}
qplot(x = residual.sugar, data = WhiteWines)
```
```{r}
summary(WhiteWines$residual.sugar)
```

```{r}
length(WhiteWines[(WhiteWines$residual.sugar < 1),]$residual.sugar)
length(WhiteWines[(WhiteWines$residual.sugar > 45),]$residual.sugar)
```

We also got a left skewed distribution with most of the data concentrated in less than 20 g/L. With the less at 0.60 and highest at 65.800 g/L and second highest at 31.60 g/L.

That been investigated we got that we have 77 less than 1 and 1 wine with residual sugar more than 45 g/L.

```{r}
WhiteWines$log10.residual.sugar <- log10( WhiteWines$residual.sugar)
qplot(x = log10.residual.sugar, data = WhiteWines)
```

```{r}
summary(WhiteWines$log10.residual.sugar)
```

I transformed the long tail data to better understang the distribution of residual sugar. The residual sugar appears bimodal with the peaks in 0.25 and 0.8.


### Total Sulfur Dioxide

Amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

```{r}
qplot(x = total.sulfur.dioxide, data = WhiteWines)
```

```{r}
summary(WhiteWines$total.sulfur.dioxide)
```

It is a normal with it peak near 130 ppm, most of the data concentrated between 0-250 ppm and the minimum at 9 ppm and max at 440ppm.


```{r}
length(WhiteWines[(WhiteWines$total.sulfur.dioxide > 50),]$total.sulfur.dioxide)/length(WhiteWines$total.sulfur.dioxide)
```

We also got that around 98.9% of the dataset have a more the SO2 more present in the wine. 

### Free Sulfur Dioxide

The free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine.

```{r}
qplot(x = free.sulfur.dioxide, data = WhiteWines)
```

```{r}
summary(WhiteWines$free.sulfur.dioxide)
```

Also a normal with concentration in beetween 0 and 100 and it peak near 50 and min value 
at 2 and max 289.


Lets see the others values as well. I suppose that ph and density may be also correlated.

### PH

Describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale


```{r}
qplot(x=pH, data=WhiteWines)
```

```{r}
summary(WhiteWines$pH)
```


We get that all wines are acid (pH < 7) and it has a normal distribution most of it at 3.1
and min 2.720 and max 3.820.

### Density

The density of water is close to that of water depending on the percent alcohol and sugar content.

```{r}
qplot(x=density,data=WhiteWines)
```


```{r}
summary(WhiteWines$density)
```

It is a normal peak at of it at 0.9937 and min at 0.987 and max near 1.04. It is concentrated at 0.9 to 1.00. That means that density of most wines are very near to the water density, that is 1.

### Chlorides

Is the amount of salt in the wine. I wonder how this can affect quality.

```{r}
qplot(x=chlorides,data=WhiteWines)
```


```{r}
summary(WhiteWines$chlorides)
```

Is also a normal with peak at 0.043 and it is concentrated between  0.009 and 0,09 and the min is at 0.009 and max at 0.34. I suppose  be the less than 0.04 have best quality,


### Sulphates

A wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant


```{r}
qplot(x=sulphates,data=WhiteWines)
```


```{r}
summary(WhiteWines$sulphates)
```

We got a binormal with peaks near 0.4 and 0.5. Min value is 0.2 and max 1.08.


### Acids

Lwts see the acid. There are three types of acids:
   1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
   2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
    3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
    
    
```{r}
qplot(x=fixed.acidity,data=WhiteWines)
```

```{r}
summary(WhiteWines$fixed.acidity)
```


```{r}
qplot(x=volatile.acidity,data=WhiteWines)
```
```{r}
summary(WhiteWines$volatile.acidity)
```

```{r}
qplot(x=citric.acid,data=WhiteWines)
```

```{r}
summary(WhiteWines$citric.acid)
```

Comparasion:

In all of them we got a normal. We have more fixed acid than the others and it peak is in 7 when the other is in 0.3 and with a higher variance.

Lets aggregate the acids and see what it returns.

```{r}
WhiteWines$acid <- WhiteWines$citric.acid + WhiteWines$fixed.acidity + WhiteWines$volatile.acidity
```

```{r}
qplot(x=acid,data=WhiteWines)
```

```{r}
summary(WhiteWines$acid)
```

When analyzing the new feature we got that it is more similar to fixed aciity with it peak at 7.5.

### Univariate Analysis

### What is the structure of your dataset?

TODO:

In my dataset there are 4898 values with 13 features (quality, residual.sugar,sulphate,fixed.acid,t).

From our observations:

- quality: There were more wines with quality equal to 6.
- alcohol: Alcohol have a binormal distribution peaks on 9% and 11% 
- density: Nomal distribution concentrated betwwen 0.99 to 1.00.
- residual sugar: Got a left skewed distribution than transformed to log10 and got a binormal with peaks at 0.25 and 0.9.
- ph: All wines are acids and had a normal distriution and peak at 3.2 
- chlorides: A normal distribution and concentrated beetween 0.02 to 0.09

### What is/are the main feature(s) of interest in your dataset?

TODO: Make a search.

The main features in the dataset are quality,alcohol and density I'd like to determine which features are better to predict the quality of a wine. I suspect that a combination of density, alcohol and others features can contribute for it.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Residual sugar,chlorides,ph,sulphates and others my contribute to the quality of a wine. 

### Did you create any new variables from existing variables in the dataset?

I created three differente features one for factorizing the quality of the wine, other by transfoming the residual sugar feature to log10 and other by combining the three different acids (fixed, volative and citric). I did so in other to better visualize the relations between the differents features.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

The residual sugar had a left skewed plot so I tried the log10 transformation and got a binormal distribution.


# Bivariate Plots Section

We will now see the correlation matrix. To see wich variables are correlated to quality and each other.

```{r}
correlation_matrix <- cor(WhiteWines[,c(1:13,15,16)])
round(correlation_matrix,2)
```

Looking more carefully with the row quality we got that quality is positivly and more correlated to alcohol. Other feature that is correlated but negatively is with density and clorides.

```{r}
sort(correlation_matrix[13,])
```

Lets  see a visualization of this matrix.

```{r echo=FALSE, Bivariate_Plots}
library(psych)
pairs.panels(WhiteWines, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```


# Bivariate Analysis

Analyzing the covariation map we got that most important ones are:

  - density and residual sugar (0.84)
  - alcohol and density (0.78)
  - total sulfur dioxide and density (0.62)
  - Quality and alcohol (0.44)
  - Quality and density (-0.33)
  

### Density and residual sugar
  
  We got by a correlation of 0.84
  
```{r}
ggplot(aes(y = density,x = residual.sugar),data = WhiteWines) + geom_point(alpha=0.5, size=1,position='jitter') + geom_smooth(method = 'lm')
```

We got almost like a linear correlation between the residual sugar and density points. That shows us that as density increases the residual.sugar also increase.

## Density and Alcohol

```{r}
ggplot(aes(y = density,x = alcohol),data = WhiteWines) + geom_point(alpha=0.5, size=1,position='jitter') + geom_smooth(method = 'lm') + scale_y_continuous(limits = c(0.98,1.04))
```

We can see that as alcohol increases there are less density. We can also get that the range of density is from 0.98 to 1.04 and alcohol range goes from 8 to 14.

### total sulfur dioxide and density

```{r}
ggplot(aes(y = density, x = total.sulfur.dioxide),data = WhiteWines) + geom_point(alpha=0.5, size=1,position='jitter') + geom_smooth(method = 'lm') + scale_y_continuous(limits = c(0.98,1.04))
```

We got that  alsmost as density increases the total sulfur dioxide also increase.

## Quality Multivariate plots

In this part we will create a plot that involves the quality of wines.

### Quality and Alcohol

Lets take a closer look at the relation between alcohol and density with quality.

```{r}
ggplot(aes(y = alcohol,x = QualityCategory),data = WhiteWines) + geom_boxplot() + geom_smooth(method = 'lm')
```

```{r}
tapply(WhiteWines$alcohol,WhiteWines$Quality, summary)
```

We see that as alcohol in the wines increases the quality tend to increase as well and that with alcohol percentage more than 10.8 tend to have better quality.
Now lets take a look at density

### Quality and density

Density is another feature that may have a correlation to quality. Lets analyse that correlation.

```{r}
ggplot(aes(x = QualityCategory, y = density),data = WhiteWines) + geom_boxplot()
```


```{r}
tapply(WhiteWines$density,WhiteWines$Quality, summary)
```

It seems that low density tend to a better quality. As it goes to near 0.99 it's Quality tend to increases.


### Quality and Clorides

According to the matrix of correlation clorides also have a correlation with quality.

```{r}
ggplot(aes(x = QualityCategory, y = chlorides),data = WhiteWines) + geom_boxplot()
```


```{r}
tapply(WhiteWines$chlorides,WhiteWines$Quality, summary)
```

As it shown in density, the clorides as it get lower than 0.02 it tend to have better quality.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

I noticed some relationships with quality of wines that are higher alcohol, lower density and loweer clorides tend to have a better quality.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

I noticed some relationships in the dataset. Density is increases proportionaly to residual sugar and sulfur dioxide but inverse proportionaly to alcohol. 

### What was the strongest relationship you found?

We got that as the density increases so do residual sugar and sulfur dioxide but inverse proportinaly to alcohol.

# Multivariate Plots Section


```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(y = density,x = alcohol),data = WhiteWines) + geom_point(aes(color=QualityCategory),alpha=0.5, size=1,position='jitter') + geom_smooth(method = 'lm') + scale_y_continuous(limits = c(0.985,1.00))
```


```{r echo=FALSE}
p <- ggplot(aes(y = density,x = alcohol,color= QualityCategory),data = WhiteWines) + geom_point(alpha=0.5, size=0.2,position='jitter') + scale_y_continuous(limits = c(0.985,1.00))
p + facet_grid(QualityCategory ~ ., scales="free")+ ggtitle('Density by Alcohol in differents Qualities')
```

In this plot you can see that as the quality increases there are more alcohol at higher levels and lower densities.


# Alcohol and cholrides
```{r echo=FALSE}
p <- ggplot(aes(y = chlorides,x = alcohol,color= QualityCategory),data = WhiteWines) + geom_point(alpha=0.5, size=0.2,position='jitter')
p + facet_grid(QualityCategory ~ ., scales="free")+ ggtitle('Chlorides  by Alcohol in differents Qualities')
```

As shown with the density, as the clhorides increases there are more alcohol at higher levels and lower densities.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
### Were there any interesting or surprising interactions between features?

The features that are strong related are alchol and quality and inversaly related to chlorides.


### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
Category plot
```{r echo=FALSE, Plot_One}
WineQuality <- table(WhiteWines$quality)
barplot(WineQuality, main="Wine Quality Distribution", xlab="Number of Wines")

```

### Description One

The wines were graded from 0 to 10 and the minimum got 3 and maximum 9. The distribution also got a shape of a normal with it peak at 6.

### Plot Two
Category and Alcohol plot
```{r echo=FALSE, Plot_Two}
ggplot(aes(y = alcohol,x = QualityCategory),data = WhiteWines) + geom_boxplot() + geom_smooth(method = 'lm')+ 
  ggtitle("Distribution of alchohol percentage by wine quality.")

```


### Description Two

We got that it has a strong relationship beetween Quality of a Wine and alchol.As alcohol percentage increases so tend to increase the quality. Although the relationship is not 100% take level 5 per exmple where we have a mean leass than the 4 level of quality.


### Plot Three
Category and Alcohol and Density plot
```{r echo=FALSE, Plot_Three}
p <- ggplot(aes(y = density,x = alcohol,color= QualityCategory),data = WhiteWines) + geom_point(alpha=0.5, size=0.2,position='jitter') + scale_y_continuous(limits = c(0.985,1.00))
p + facet_grid(QualityCategory ~ ., scales="free")+ ggtitle('Density by Alcohol in differents Qualities')
```

### Description Three
In this graph we compare two,density and alchol, variables correlated to quality of the wine.
In this plot you can see that as the quality increases there are more alcohol at higher levels and lower densities.
------

# Reflection

We got that features that are strong related to wine are alcohol, density and chlorides. Quality is proportinaly related to alcohol and inverse to density and chlorides. 
A suggestion to inprove the dataset will be to take into account wines from differents parts of the world. In this case of study were only considered wines from Portugal. By doing so, we can remove some bias.


